Spin-orbit misalignment for the long period planet candidate KOI-368.01

### 2013-07-04 msV0.tex
mergied and has the draft to send to Gaspar

\documentclass[apjl]{emulateapj}

\shorttitle{Spin-orbit misalignment for KOI-368}
\shortauthors{shortauthors}

%%% packages
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xspace}
%\usepackage{array}
\usepackage{lineno}

@@

@@

[APO observation and spectrum fitting]

The host star rotation period can be identified by searching for spot
modulation in the Kepler long cadence PDC lightcurves. We masked out
the primary transits and performed a Lomb-Scargle periodogram \citep[][]{Lomb1976,Scargle1982} analysis
on the lightcurves. A significant peak at 1.19 days was identified,
and checked by the CLEAN algorithm \citep{Roberts1987}. [This peak corresponds with vsini
and was adopted as the rotation period of the star].

@@

\section{Orbit obliquity from lightcurve asymmetry}
\label{sec:transit-light-curve}

\subsection{Lightcurve model}
\label{sec:lightcurve-model}

\begin{figure}
  \centering
  \includegraphics[width=9cm]{obliq_model.eps}
  \caption{f=0.2,beta=1.0,phi=0. Model using koi-368 system
    parameters (inc, rsum, rratio etc). 0, 45, 90 degree obliquity.}
  \label{fig:obliqmodel}
\end{figure}

@@

The transit lightcurve is modelled using the
\citet{Nelson1972} model, implemented in an adaption of the
JKTEBOP code \citep{Proper1981,Southworth2004}. The
relevant free parameters are orbital period $P$, transit centre $T_0$,
normalised radius sum $R_\star+R_p / a$, radius ratio $R_p/R_\star$,
line of sight inclination $i$. Quadratic limb darkening coefficients,
$c_1$ and $c_2$, 
are fitted for in an initial minimisation routine, with initial
estimates taken from \citet{Sing2010}, then held fixed for subsequent analyses. Jump
parameters for the stellar oblation correction include the planet
orbit obliquity $\lambda$, stellar oblation $f$. The projection angle
between the stellar rotation axis and line of sight $i_\text{rot}$ is
fixed for an initial analysis, then set free to explore the potential
degeneracies. The gravity darkening exponent $\beta$ is set to be
fixed at 0.25 [cite], but also allowed free in subsequent analysis to
explore degeneracies. A flux offset for each transit event is calculated and
removed at each iteration, and is not included in the fit
parameters. For Kepler long cadence data, the model is modified by a
30 minute boxcar smooth. The best fit parameters and the posterior
probability distribution is explored via a Markov chain Monte Carlo
(MCMC) analysis, using the \emph{emcee} MCMC ensemble sampler
\citep{ForemanMackey2012}. The likelihood function is given by
$\exp(-\Delta\chi^2/2)$. For each transit, we scale the flux errors such
that the reduced $\chi^2$ is at unity. This allows for errors other
than photon noise to be taken into account.

Figure~\ref{fig:lightcurve} plots the phase folded transit lightcurve of
KOI-368.01 and the best fit spherical and oblate models. The best fit
spherical host star model cannot explain the significant in-transit
asymmetry observed. The best fit parameters are presented in Table~\ref{tab:params}.

The posterior probability distributions for relevant parameters
are plotted in Figure~\ref{fig:posterior}. We find a weak inverse dependency
between $\lambda$ and $\beta$, and a positive dependency between
$\lambda$ and $R_p/R_\star$.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=7cm]{ecc_model.eps}
  \caption{The maximum lightcurve distortions due to a
    eccentric orbit for KOI-368.01 are plotted as residuals to its
    best fit circular orbit model. The maximum velocity change is
    determined by measuring the difference in the ingress and egress
    duration. Eccentricity cannot explain the observe lightcurve
    asymmetry (Figure~\ref{fig:lightcurve}, middle panel).}
  \label{fig:eccmodel}
\end{figure}

@@

[Blends in images]

@@

@@

@@

>>>>

\documentclass[preprint]{emulateapj}
\input{newcommand_gen.tex}
\shorttitle{Spin-orbit misalignment for KOI-368}
\shortauthors{shortauthors}

%%% packages
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{subfigure}
%\usepackage{array}
\usepackage{lineno}

@@

\citet{Hirano:2012} 
using the rotation period from the \lc\ and 
then combined with the projected rotation period and stellar radius from 
high-res spectroscopy, measured the obliquity of 15 KOI systems.

\citet{Barnes:2011} use the asymmetry in the Kepler transit \lc\ of 
KOI-13 to measure the spin-orbital alignment angle.

%\section{Method}
%\label{sec:method}

%The gravity darkening model

@@

{\bf KOI-368} is assigned as an Kepler Object of Interesting (KOI) in the 
\citet{Batalha:2013} Kepler catalog from the Quarter 1-6 data. It has a 
Kepler Input Catalog (KIC) entry 6603043. The KIC parameters of the star are 
$T_{\rm eff} = 9034 {\rm K}$, $\rm logg=4.133$, and $\rm FeH = -0.028$, 
indicating it is a hot and fast rotating star.

We took an high resolution (R$\sim$31500) spectrum of this star using 
the ARC Echelle Spectrograph mounted on the Apache Point Observatory  
3.5-meter telescope (APO 3.5m).
[APO observation and spectrum fitting]

The host star rotation period can be identified by searching for spot
modulation in the Kepler Long Cadence PDC \lcs\ \citep{Smith:2012} 
from Q1-Q9. We masked out the primary transits and performed a Lomb-Scargle 
periodogram \citep[][]{Lomb1976,Scargle1982} analysis
on the \lcs. A significant peak at 1.19 days was identified,
and checked by the CLEAN algorithm \citep{Roberts1987}. [This peak 
corresponds with vsini and was adopted as the rotation period of the star].

@@

\section{Orbit obliquity from light curve asymmetry}
\label{sec:transit-light-curve}

\subsection{Light curve model}
\label{sec:lightcurve-model}

We make use of the all the available public Kepler \lcs\ for our analysis 
of the photometry effect predicted by \citet{Barnes:2009}. These include 
Long Cadence (29.4min) data (Q0-Q15) and Short Cadence (58.84s) data (Q8-Q9). 
The long cadence \lcs\ consist of 11 transit altogether (assigned 
transit number from epoch 0-11, the 7th transit is not observed by Kepler 
due to data gap). The short cadence \lcs\ have transit 5 and 6.

We start with the $\rm SAP\_FLUX$ obtained from the MAST arxive  
\footnote{http://archive.stsci.edu/kepler/data$\_$search/search.php}, the out 
of transit variations are corrected by the following steps from 
\citet{Huang:2013}:

a) removal of bad data points;

b) correction of safe mode and tweaks;

c) a set of cosine functions with minimum period of 1 day.

d) a 7th order polynomial fit on the out of transit parts. 
\begin{figure}
  \centering
  \includegraphics[width=9cm]{obliq_model.eps}
  \caption{f=0.2,$\beta=1.0$, $i_{\rm rot}=0$. Model using koi-368 system
    parameters (inc, rsum, rratio etc). 0, 45, 90 degree obliquity.}
  \label{fig:obliqmodel}
\end{figure}

%\begin{figure}[h!]
%  \centering
%  \includegraphics[width=7cm]{LS.eps}
%  \caption{Lomb-Scargle periodogram of the PDC lightcurve, with the
%    transits masked out. The lightcurve is rotationally modulated with
%  a period of 1.19 days.}
%  \label{fig:LS}
%\end{figure}

\subsection{Gravity darkening modeling}
\label{sec:grav-model}
The gravity darkening model is generated following \citet{Barnes:2009}.
The temperature profile on the stellar surface is determined by the local 
effective gravity \citep{Zeipel:1924}. 
\begin{equation}
\frac{T}{T_{\rm pole}} \propto (\frac{g_{\rm eff}}{g_{\rm pole}})^{\beta} 
= F_{f,w} (r,\theta).
\end{equation}
The the effective gravity is defined as 
\begin{equation}
\vec{g}_{\rm eff} = -\Omega_{\rm grav}^2\frac{R_{\rm eq}^3}{R^2}\hat{r} \\ 
+ \Omega_{ror}^2\,R_{\perp}\hat{r_{\perp}}
\end{equation}
We use $\Omega_{\rm ror}$ to measure the stellar rotation rate in radians 
per second and $\Omega_{\rm grav}$ to represent the angular velocity due to 
gravity at the equator, $\Omega_{\rm grav}^2 = GM/R_{\rm eq}^3.$). 
The definition of the other symbols follow \citet{Barnes:2009} Eq.10, in 
which $G$ is the gravitational constant, $M$ is the mass of the star, $R$ 
and $R_{\perp}$ are the distance from the stellar center and stellar 
rotation axis to the point of question, respectively. $\hat{r}$ and 
$\hat{r_{\perp}}$ are unit vectors indicate the directions of $R$ and 
$R_{\perp}$. 
The effective gravity profile $\frac{g_{\rm eff}}{g_{\rm pole}}$ at the stellar 
surface is a function of the oblation of the star $f$, defined as 
$\frac{R_{\rm eq}-R{\rm pole}}{R_{\rm eq}}$, the non-dimension 
measure of the rotation rate ($w=\Omega_{ror}/\Omega_{grav}$, 
and the normalized position parameters $r$ and $\theta$. The 
We use the gravity darkening coefficient computed 
for Kepler band using ATLAS code from \citet{Claret:2011}. By assuming the 
distribution of errors in the stellar parameters to be gaussian, we obtain 
the gravity darkening parameter $\beta$ to be ?.
We demonstrate the flux profile computed for KOI-368 system with $f=0.2$, 
$\beta=1.0$ and assuming the stellar rotation axis is in the plane 
perpendicular to our line of site ($i_{\rm rot}=0$) in Figure 
\ref{fig:obliqmodel}. If the planet cross the hotter pole to colder 
equator during its transits, the transit \lcs\ will show asymmetry depend 
on the misalignment between the planet orbit and stellar rotation axis.

@@

The transit \lcs\ are modeled using the
\citet{Nelson1972} model, implemented in an adaption of the
JKTEBOP code \citep{Proper1981,Southworth2004}. The
relevant free parameters are orbital period $P$, transit centre $T_0$,
normalised radius sum $R_\star+R_p / a$, radius ratio $R_p/R_\star$,
line of sight inclination $i$. Quadratic limb darkening coefficients,
$c_1$ and $c_2$, are fitted for in an initial minimization routine, with 
initial estimates taken from \citet{Sing2010}, then held fixed for 
subsequent analyses. Jump parameters for the stellar oblation correction 
include the planet orbit obliquity $\lambda$, stellar oblation $f$. The 
projection angle between the stellar rotation axis and line of sight 
$i_\text{rot}$ is fixed for an initial analysis, then set free to explore 
the potential degeneracies. The gravity darkening exponent $\beta$ is set 
to be fixed at (? see Section \S \ref{sec:grav-model}), but also allowed 
free in subsequent analysis to explore degeneracies. A flux offset for each 
transit event is calculated and removed at each iteration, and is not 
included in the fit parameters. For Kepler long cadence data, the model is 
modified by a 30 minute boxcar smooth. The best fit parameters and the 
posterior probability distribution is explored via a Markov chain Monte Carlo
(MCMC) analysis, using the \emph{emcee} MCMC ensemble sampler
\citep{ForemanMackey2012}. The likelihood function is given by
$\exp(-\Delta\chi^2/2)$. For each transit, we scale the flux errors such
that the reduced $\chi^2$ is at unity. This allows for errors other
than photon noise to be taken into account.

Figure~\ref{fig:lightcurve} plots the phase folded transit lightcurve of
KOI-368.01 and the best fit spherical and oblate models. The best fit
spherical host star model cannot explain the significant in-transit
asymmetry observed.

[compare with Kepler original parameters]
The transit parameter from Kepler catalog is 
$P = 110.3216019$ day and $R=19.16$\rearth.

[restate the finding of oblation]
We find the transit \lcs\ are more consistent with a model of ...

The best fit parameters are presented in Table~\ref{tab:params}.
The posterior probability distributions for relevant parameters
are plotted in Figure~\ref{fig:posterior}. We find a weak inverse dependency
between $\lambda$ and $\beta$, and a positive dependency between
$\lambda$ and $R_p/R_\star$.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=7cm]{ecc_model.eps}
  \caption{The maximum lightcurve distortions due to a
    eccentric orbit for KOI-368.01 are plotted as residuals to its
    best fit circular orbit model. The maximum velocity change is
    determined by measuring the difference in the ingress and egress
    duration. Eccentricity cannot explain the observe lightcurve
    asymmetry (Figure~\ref{fig:lightcurve}, middle panel).}
  \label{fig:ecc_model}
\end{figure}

@@

To establish the planetary nature of KOI-368.01, we investigate in the 
false positive scenarios of this system. We use the image from APO 3.5m 
Echelle Slitviewer camera to rule out close by bright companions. The 
camera has a field of view of 63.6$\arcsec$ and the pixel scale is 0.133\pxs. 
We can rule out companion brighter than 20 mag within 20\sqarcsec\ 
area of KOI-368 (See Figure \ref{fig:APO}). No candidate is found to be 
a possible source of background blend.

We are able to use the upper limit of centroid displacement to rule out the 
possibility that the transit signal is mimicked by background binaries. We 
use the flux weighted first momentum centroids produced by Kepler 
pipeline. The long time range trend in centroids due to the motion of 
the space craft is corrected by a 7-th order polynomial on each quarter 
of the Long Cadence data. For each Long Cadence transit [do we want to put 
the SC points on as well?], we compare the centroids of star during the 
transit with the centroids before and after the transit (within a window 
of width the same as the transit duration). The displacements between the 
mean centroids of the in-transit part and the out-of-transit part are shown 
in Figure \ref{fig:centroid}. The error-bars are computed with the error 
propagation rule by assuming the centroids in each window follows a 
Gaussian distribution. The analysis shows that the signal could not be 
from a source farther than 1$\arcsec$ away from the star KOI-368.

@@

The Transit Timing Variation (TTV) analysis for KOI-368.01 with the 
11 transits from Long Cadence data prefer a solution without any high 
amplitude TTV signal (See Figure \ref{fig:TTV}).

@@

\begin{figure}
\subfigure[]
{
\includegraphics[width=0.4\linewidth]{APOimage_large.eps}
\label{fig:APO}
}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{centroid.eps}
\label{fig:centroid}
}
\caption{
caption
\label{fig:FPs}
}
\end{figure}

@@

We measured the asymmetry in the transit lightcurve of KOI-368.01, and 
found it to be orbiting in a highly inclined plane of 
$\lambda = 50\pm20^\circ$. With a period of 110.3 days, it is one of the 
longest period planets for which spin-orbit alignment has been measured. 
HD80606b, the only other long period single planetary system 
with obliquity measured, has almost identical period and projected orbit 
obliquity, is highly eccentric $e=0.93$. The high eccentrcity and obliquity 
of the HD80606b system can be explained satisfily with Kozai cycle excited 
by its main-sequence companion 1000 AU away \citep{WuMurray:2003}. 
Curiously, we constrain the eccentricity of the KOI-368 system to be much 
lower, $e = 0.07^{+0.18}_{-0.07}$. If Kozai cycle is responsible for this 
high obliquity, the low eccentricity put it in a "quenched" stage of the 
migration. Following Equation 12 of \citet{Socrates:2012}, this is 
unlikely be caused by solar type companian within 100 AU or stellar type 
companian within 50 AU. 
But the age of the system (given the host is an A star 
and an age indication from the spectrum analysis) is much younger than 
the typical eccentricity dumping timescale (2-3 Gyrs).

In particular, the oblique orbit of KOI-368.01 ... 
[important for migration theory X]

\begin{figure}
\includegraphics[width=0.9\linewidth]{TTV.eps}
\caption{caption
\label{fig:TTV}}
\end{figure}

### 2013-07-04 msV1.tex
mergied and has the draft to send to Gaspar

\begin{figure}
\subfigure[]
{
\includegraphics[width=0.4\linewidth]{APOimage_large.eps}
\label{fig:APO}
}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{centroid.eps}
\label{fig:centroid}
}
\caption{
caption
\label{fig:FPs}
}
\end{figure}

\section{Discussion}
\label{sec:discussion}

We measured the asymmetry in the transit lightcurve of KOI-368.01, and 
found it to be orbiting in a highly inclined plane of 
$\lambda = 50\pm20^\circ$. With a period of 110.3 days, it is one of the 
longest period planets for which spin-orbit alignment has been measured. 
HD80606b, the only other long period single planetary system 
with obliquity measured, has almost identical period and projected orbit 
obliquity, is highly eccentric $e=0.93$. The high eccentrcity and obliquity 
of the HD80606b system can be explained satisfily with Kozai cycle excited 
by its main-sequence companion 1000 AU away \citep{WuMurray:2003}. 
Curiously, we constrain the eccentricity of the KOI-368 system to be much 
lower, $e = 0.07^{+0.18}_{-0.07}$. If Kozai cycle is responsible for this 
high obliquity, the low eccentricity put it in a "quenched" stage of the 
migration. Following Equation 12 of \citet{Socrates:2012}, this is 
unlikely be caused by solar type companian within 100 AU or stellar type 
companian within 50 AU. 
But the age of the system (given the host is an A star 
and an age indication from the spectrum analysis) is much younger than 
the typical eccentricity dumping timescale (2-3 Gyrs).

In particular, the oblique orbit of KOI-368.01 ... 
[important for migration theory X]

\begin{figure}
\includegraphics[width=0.9\linewidth]{TTV.eps}
\caption{caption
\label{fig:TTV}}
\end{figure}

>>>>

\begin{figure}
\subfigure[]
{
\includegraphics[width=0.45\linewidth]{APOimage.eps}
\label{fig:APO}
}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{centroid.eps}
\label{fig:centroid}
}
\end{figure}
\section{Discussion}
\label{sec:discussion}

\begin{figure}
\includegraphics[width=0.9\linewidth]{TTV.eps}
\caption{caption}
\label{fig:TTV}
\end{figure}
